### 2023-02-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@b99dd7488628b3c5d215763948b80c5c39ca193c 🚀

The uncertainty and spatiotemporal variability of both types of trophic
interactions (Boolean and weighted) can be represented
probabilistically. On one hand, Boolean interactions follow a Bernoulli
distribution \(A_{i,j} \sim {\rm Bernoulli}(p)\), with \(p\) being the
probability of interactions. The only two possible outcomes are the
presence (\(A_{i,j} = 1\)) or absence (\(A_{i,j} = 0\)) of an
interaction between the two nodes. Weighted interactions, on the other
hand, can follow various probability distributions depending on the
measure used. In this case, the event's outcome is the value of
interaction strength. For instance, weights can follow a Poisson
distribution \(A_{i,j} \sim {\rm Poisson}(\lambda)\) when predicting
frequencies of interactions between pairs of nodes, with \(\lambda\)
being the expected rate of interaction. Note that weighted interactions
can be converted to probabilistic interactions by normalizing. The
definition and interpretation of parameters like \(p\) and \(\lambda\)
are inextricably linked to environmental and biological factors such as
species relative abundance, traits, area, and time, depending on the
type of interaction.\\
Because Boolean species interactions are much more documented in the
literature, our primary focus in this contribution will be on addressing
the challenges in defining and interpretating \(p\) for pairwise species
interactions.

@@

\hypertarget{exploring-different-levels-of-organization}{%
\subsection{Exploring different levels of
organization}\label{exploring-different-levels-of-organization}}

How do interaction probabilities scale taxonomically?

\begin{itemize}
\tightlist
\item
  There are different biological interpretations of probabilities in
  food webs at the individual level and at higher taxonomic levels.
\item
  How does the scaling up of the nodes from an individual to population
  to any higher taxonomic group change our interpretation of interaction
  probabilities? How does the aggregation change our interpretation?
\item
  Why would we want to scale networks taxonomically?
\item
  Do all nodes need to be the same taxonomic scale, within a network?
\item
  How is it similar and different to spatial and temporal scaling?
  Basically, all kinds of scaling are just different ways to aggregate
  individuals or nodes.
\item
  Papers: Vázquez \emph{et al.} (2022)
\end{itemize}

>>>>

The uncertainty and spatiotemporal variability of both types of trophic
interactions (Boolean and weighted) can be represented
probabilistically. On one hand, Boolean interactions follow a Bernoulli
distribution \(A_{i,j} \sim {\rm Bernoulli}(p)\), with \(p\) being the
probability of interactions. The only two possible outcomes are the
presence (\(A_{i,j} = 1\)) or absence (\(A_{i,j} = 0\)) of an
interaction between the two nodes. Weighted interactions, on the other
hand, can follow various probability distributions depending on the
measure used. In this case, the event's outcome is the value of
interaction strength. For instance, weights can follow a Poisson
distribution \(A_{i,j} \sim {\rm Poisson}(\lambda)\) when predicting
frequencies of interactions between pairs of nodes, with \(\lambda\)
being the expected rate of interaction. Note that weighted interactions
can be converted to probabilistic interactions by normalizing. The
definition and interpretation of parameters like \(p\) and \(\lambda\)
are inextricably linked to environmental and biological factors such as
species relative abundance, traits, area, and time, depending on the
type of interaction. Because Boolean species interactions are much more
documented in the literature, our primary focus in this contribution
will be on addressing the challenges in defining and interpretating
\(p\) for pairwise species interactions.

@@

\hypertarget{exploring-different-taxonomic-resolutions}{%
\subsection{Exploring different taxonomic
resolutions}\label{exploring-different-taxonomic-resolutions}}

The properties of ecological networks depend on their level of
organization (Guimarães 2020). Indeed, at different taxonomic scales,
different behaviours and dynamics can be observed and distinct
ecological questions can be answered (e.g., exploring evolutionary
dynamics at broad taxonomic scales). Because of these reasons, it could
be important to analyse the same network at different taxonomic scales.
However, we want to emphasize here that many networks do not have an
homogenous level of organisation (Vázquez \emph{et al.} 2022). Indeed,
different nodes within the same network can be represented at different
taxonomic scales (e.g., a network composed of species and trophic
species). This becomes important when we consider that the biological
interpretation of interaction probabilities depends on the nodes'
resolution. For example, in individual-based networks, the probability
that two individuals interact could represent the degree of belief that
one will actually consume the other. In species-based networks, the
probability that two species interact could rather represent the degree
of belief that \emph{at least} one individual from the predator species
will eat \emph{at least} another individual from the prey species. This
distinction in interpretation impacts the way probability values change
with taxonomic scale.

There are a lot of similarities between taxonomic and spatiotemporal
scaling of probabilistic interactions. Fundamentally, these types of
scaling are just different ways to aggregate individuals into broader
nodes, either spatially, temporally, or taxonomically. However, there
are also important differences between them. First, in metawebs, if we
know that two species have the capacity to interact, we can infer that
their respective genus should also be able to interact (i.e., there
should be at least two individuals within these genus that can
interact). On the contrary, knowing that two genus can interact does not
mean that all pairwise combinations of species within these genus can
also interact among themselves. This observation also applies to local
networks. When it comes to probabilistic networks, interaction
probabilities at broader taxonomic scales can be directly obtained from
probabilities at finer scales when aggregating nodes. For example, if we
have in a network \(n_A\) species from genus \(A\) and \(n_B\) species
from genus \(B\), we can calculate the probability that the two genus
interact as
\(P_N(A \rightarrow B) = 1 - \prod_{i = 1}^{n_1}\prod_{j = 1}^{n_2}(1 - P_N(A_i \rightarrow B_j))\),
where \(A_i\) and \(B_j\) are the species of the corresponding genus.
However, more sophisticated models need to be used when building
probabilistic networks at smaller taxonomic resolutions (e.g., when
building a species-level network from a genus-level network). One could,
for example, estimate the probabilities of all pairwise species
interactions by using a Beta distribution parametrised by the
broader-scale network.
